---
title: "Week 3 Exercises"
author: "J.Blondeau"
date: "2023-09-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("tidyverse")
library(tidyverse)

```

All the following question use the sttst_all_sites.xlsx file in the /data directory.

All answers should be made in the .Rmd file, showing the code and your answer and everything pushed back up to your GitHub repo.

triple stars renders a thin line across the page

***
1. Import sttstj_all_sites and inspect the data set.  What is the data structure and the data type of each field?

Data structure is tabular. data types for yr_site and hab are characters and lat and lon are numeric. 

```{r}
 #this is a code chunk, anything inside the back tics is like being in the R console.
 
#data structure is tabular. data types for yr_site and hab are characters and lat and lon are numeric. 


```
***

2. How many samples are in the data set?   

2505 samples. 

```{r}
#2505 samples 


```

3. How many unique habitat types are there?  What are the names of each habitat?

There are 6 unique habitat types. PVMT, AGRF, SCR, PTRF, BDRK, HARD.

```{r}

unique(sttstj_all_sites$hab)

#there are 6 unique habitat types. PVMT, AGRF, SCR, PTRF, BDRK, HARD

```

4. How many sites are there in the "PVMT" habitat type?

571 sites. 

```{r}
pvmtsites <- sttstj_all_sites %>% filter(hab == "PVMT")

#there are 571 sites. 

```

5. How many sites are there in the "PVMT" or "AGRF" habitat types?

There are 1538 sites in the PVMT and AGRF habitats. Individually, there are 571 sites in PVMT and 967 sites in AGRF. 

```{r}
pvmtagrfsites <- sttstj_all_sites %>% filter(hab == "PVMT" | hab == "AGRF")
pvmtagrfsites
pvmtsites <- sttstj_all_sites %>% filter(hab == "PVMT")
agrfsites <- sttstj_all_sites %>% filter(hab == "AGRF")

#there are 571 sites in PVMT and 967 sites in AGRF.
#1538 total sites

```

6. How many sites are there in all the habitat types EXCEPT "PVMT"?

1934 sites. 

```{r}
allbutpvmtsites <- sttstj_all_sites %>% filter(hab != "PVMT")

#1934 sites

```
7. Show the first 5 records (rows) of...
    - all sites sampled in "AGRF" or "SCR"
    - only show the "yr_site" and "hab" columns
    
<!-- extra line break -->

yr_site   hab

2001_243	AGRF			
2001_244	AGRF			
2001_247	AGRF			
2001_249	SCR			
2001_256	AGRF			
2001_260	AGRF

<br>

```{r}
agrfscrsites <- sttstj_all_sites %>% filter(hab == "AGRF" | hab == "SCR") %>% select(yr_site, hab)

head(agrfscrsites)



```

8. What is the northern most site sampled in the "BDRK" habitat?

Site 5450.

```{r}
bdrksites <- sttstj_all_sites %>% filter(hab == "BDRK") %>% select(yr_site, lat, lon, hab)

View(bdrksites)

max(bdrksites$lat)

latbdrk <- bdrksites %>% arrange(desc(lat))

head(latbdrk)

#18.40479, -65.05035

#2019_5450

```

9. What is the western most site sampled in either "AGRF" or "PTRF" habitat?

Site 9758.

```{r}
agrfptrf <- sttstj_all_sites %>% filter(hab == "AGRF" | hab == "PTRF") %>% select(yr_site, lat, lon, hab)
View(agrfptrf)

agrfptrflon <- agrfptrf %>% arrange(lon)
head(agrfptrflon)

#2015_9758

#18.28121, -65.14093

```

10. How many years were surveyed in the data set?

16 years. 

```{r}
split <- sttstj_all_sites %>% separate(yr_site, c('year', 'site'))
View(split)

unique(split$year)

#16 years

```

11. What is the southern most site sample in the "PVMT" habitat in 2004?

Site 1932

```{r}
pvmtyr <- split %>% filter(year == "2004", hab == "PVMT")
View(pvmtyr)

pvmtlat <- pvmtyr %>% arrange(lat)
head(pvmtlat)

#1932


```


